### 2024-03-06 src/tex/acronymns.tex
start using showyourwork

\newacronym{llm}{LLM}{large language model}
\newacronym{ml}{ML}{machine learning}
\newacronym{agi}{AGI}{artificial general intelligence}
\newacronym{api}{API}{application programming interface}
\newacronym{mcq}{MCQ}{multiple-choice question}
\newacronym{rest}{REST}{Representational State Transfer}
\newacronym{orm}{ORM}{object relational mapping}