### 2024-03-07 src/tex/safety_data.tex
Create safety_data.tex

\subparagraph{GHS classification, hazard statements and daily allowed intakes}

The GHS classification, hazard statements and daily allowed intake (DAI) data has been extracted from PubChem. 
The former two have been mined through python, while the latter has been manually acquired.
The DAIs have been curated to contain only records approved by the World Health Organization (WHO). 
The chemicals in this class of questions belong to one of the three classes: pesticides (e.g. calcium arsenate), insecticides (e.g. cyfluthrin) or herbicides (e.g. 2,4 - D).