### 2024-03-08 src/tex/showyourwork-metadata.tex
rebuild

        \OnGithubActionsfalse
        \def\syw@url{https://github.com/lamalab-org/chembench-paper}
    \def\syw@slug{lamalab-org/chembench-paper}
    \def\syw@sha{9e3e1d5a5f3dc81ac1ab792b79e58cce26f6bbda}
    \def\syw@runid{}

        \AddStampfalse
        \newcommand{\syw@stampX}{0.5}
    \newcommand{\syw@stampY}{0.5}
    \newcommand{\syw@stampSize}{0.75}
    \newcommand{\syw@stampAngle}{-20.0}
    \newcommand{\syw@stampText}{{\faGithub}/lamalab-org/chembench-paper}
    \newcommand{\syw@stampVersion}{0.4.2}

        \addvalue{fig:topic_barplot_script}{src/scripts/plot_statistics.py}
        \addvalue{fig:wordcloud_script}{src/scripts/wordcloud.py}
        \addvalue{fig:human_timing_script}{src/scripts/analyze_human_data.py}
        \addvalue{fig:experience_vs_correctness_script}{src/scripts/analyze_human_data.py}

        \addvalue{output/total_number_of_questions.txt_rule}{Snakefile\#L1}
        \addvalue{output/manually_generated.txt_rule}{Snakefile\#L1}
        \addvalue{output/automatically_generated.txt_rule}{Snakefile\#L1}
        \addvalue{output/number_experts.txt_rule}{Snakefile\#L20}
        \addvalue{output/total_hours.txt_rule}{Snakefile\#L20}
        \addvalue{output/flesch_kincaid_reading_ease.txt_rule}{Snakefile\#L30}
        \addvalue{output/num_h_statements.txt_rule}{Snakefile\#L1}
        \addvalue{output/num_pictograms.txt_rule}{Snakefile\#L1}
        \addvalue{output/num_dai.txt_rule}{Snakefile\#L1}
        \addvalue{output/spearman_experience_score.txt_rule}{Snakefile\#L20}
        \addvalue{output/spearman_experience_score_p.txt_rule}{Snakefile\#L20}