### 2024-03-20 src/tex/parse_check_desc.tex
feat: description of parsing checks

\paragraph*{Manual validation of parsing performance}

As stated, we tried to account for the variation in outputs with custom regular expressions.
We selected a large, diverse subset of 791 questions and manually investigated where the parsed output does not match the actual answer intended by the model.
We found that for \glstext{mcq} questions, the parsing was accurate in 99.76\% of the cases, while for floating point questions, the parsing was accurate in 99.17\% of the cases.
The models generating errors are \texttt{pplx-7b-chat} and \texttt{Mixtral-8x7b}, which in the specified amount of cases fail to follow the prompt.