### 2024-03-27 src/tex/acronymns.tex
add note about leaderboard

\newacronym{llm}{LLM}{large language model}
\newacronym{ml}{ML}{machine learning}
\newacronym{agi}{AGI}{artificial general intelligence}
\newacronym{api}{API}{application programming interface}
\newacronym{mcq}{MCQ}{multiple-choice question}
\newacronym{rest}{REST}{Representational State Transfer}
\newacronym{orm}{ORM}{object relational mapping}
\newacronym{dai}{DAI}{daily allowed intake}
\newacronym{ghs}{GHS}{Globally Harmonized System of Classification and Labelling of Chemicals}
\newacronym{who}{WHO}{World Health Organization}
\newacronym{nmr}{NMR}{Nuclear Magnetic Resonance}
\newacronym{helm}{HELM}{Holistic Evaluation of Language Models}
\newacronym{smiles}{SMILES}{Simplified Molecular Input Line-Entry System}
\newacronym{pca}{PCA}{Principal Component Analysis}
\newacronym{iupac}{IUPAC}{International Union of Pure and Applied Chemistry}
\newacronym{json}{JSON}{JavaScript Object Notation}