### 2024-03-28 src/tex/sections/semi_programatically_sources_table.tex

\begin{table}[h]
\begin{tabularx}{\textwidth}{Xp{7 cm}X}
    \toprule
    source & description & question count \\
\midrule
Number of isomers & MAYGEN was used to compute the number of isomers for a set of SMILES extracted from the ZINC dataset. & \variable{output/question_count_per_dir/json_file_counts_number_of_isomers.txt} \\
Total electron count of molecules & Electron counts based on the data from https://www.cheminfo.org/. & \variable{output/question_count_per_dir/json_file_counts_electron_counts.txt} \\
Oxidation states & Oxidation states questions based on the data from https://www.cheminfo.org/. & \variable{output/question_count_per_dir/json_file_counts_oxidation_states.txt} \\
Chemical reactivity & Questions are framed based on the information from Cameo Chemicals website (https://cameochemicals.noaa.gov/reactivity). & \variable{output/question_count_per_dir/json_file_counts_reactive_groups.txt} \\
Number of NMR signals & Molecules are sampled from the ZINC database and the number of NMR signals is computed with OpenChemLib & \variable{output/question_count_per_dir/json_file_counts_number_of_nmr_peaks.txt} \\
Point group of molecules & Our ChemCaption tool is used to assign the point group, and then each case was manually checked to select well defined cases. & \variable{output/question_count_per_dir/json_file_counts_point_group.txt} \\
IUPAC-SMILES pairs & Sampled from the PubChem database & \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} + \variable{output/question_count_per_dir/json_file_counts_smiles_to_name.txt} \\
\multirow{3}{*}{PubChem safety data} & Daily allowable intakes according to the World Health Organization & \variable{output/question_count_per_dir/json_file_counts_dai.txt}  \\
 & Definitions of hazard statements &  \variable{output/question_count_per_dir/json_file_counts_h_statements.txt} \\
 & GHS classification of chemicals mined through the API & \variable{output/question_count_per_dir/json_file_counts_pictograms.txt} \\
\multirow{2}{*}{Safety}
& Materials' compatibility & \variable{output/question_count_per_dir/json_file_counts_materials_compatibility.txt} \\
 & Chemical compatibility & \variable{output/question_count_per_dir/json_file_counts_chem_chem_comp.txt} \\
\bottomrule
\end{tabularx}
\end{table}