### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
small fix

The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).
Models also struggle with identifying the relevant chemical trends (e.g., ordering the entropy of compounds).

\gls{llm} also struggle with open-ended questions. Albeit in our study, we only investigated numerical open-ended questions, thus a part of these would require multi-step calculations, which can be challenging for \gls{llm}.