### 2024-10-29 src/tex/sections/patterns_in_wrong_questions.tex
fix wording

The models seem to not be able to answer questions that require specific knowledge (e.g., safety profile of chemicals) and reasoning (e.g., oxidation number of a specific element, the number of chiral centers in an organic molecule or identifying a symmetry group).

\glspl{llm} also struggle with questions that do are not \glspl{mcq}.
Albeit in our study, we only investigated numerical outputs that can be parsed and scored. Thus, a part of these would involve multi-step calculations, which can be challenging for \glspl{llm}.\autocite{zhou2023llm_arithmetics}